---
title: "Options"
output: 
  pdf_document:
    extra_dependencies: ["bm"]
date: "2024-03-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Most of these will be option pricing algorithms. 

# Binomial Lattice
Consider the $N-$period binomial model with $0<d<e^{r\Delta t} < u$. Suppose the derivative payout at maturity $V_N$ is a random variable with known distribution. 

Idea: Starting from the leaves (known $V_N$), we will recursiveley go down levels $n= N-1, \dots, 0$ computing discounted expected values. 

* Algorithm 2.1: If we wish to price **path-dependent** options, we must keep track of **exact** path evolution at each time $n$:$(\omega_1, \dots, \omega_n)$ 
\[ V_n(\omega_1, \dots, \omega_n) = e^{-r\Delta T}\mathbb{E}^{\mathbb{Q}}(V_{n+1}|\mathcal{F}_n) \]

* Algorithm 2.2: To price **path-independent** option, we can simplify Algorithm 2.1 and just count the number of ups. This uses the fact that $V_t^j$ is the same in the binomial lattice no matter independent of the path and only dependent on the number of ups, ie. trajectories $(up, down, up), (down, up, up)$ and $(up,up,down)$ have the same stock value $V_3^2$.


## Initialization
1. Given $S_0, d, u, p^u$ we will compute the lattice evolution up to time $N$ where:
\[ S_{t}^j = S_0u^jd^{t-j}\]
for all $t=1,2,\dots, N$, $j=0,\dots, t$. Note that $S_t^j$ stands for "stock price at time $t$ with state $j-$ups from t=0"


## Algorithm 2.1
To find the price of any option

2. Compute all possible payouts at maturity. $V_T(\omega_1, \dots\omega_n)$ for $\omega_1, \dots\omega_n \in \Omega=\{up, down\}$

3. For $n=N-1, \dots, 0$ do: for all $2^n$ states $W = \omega_1, \dots \omega_n$ compute:

\[ V_n(W)= e^{-r \Delta t}(q^uV_{n+1} (W, up)+ q^d V_{n+1}(W, down)) \]

4. Return $V_0$

This has a complexity of $2^T$ when implemented in a naive way because it goes through all possible $2^T$ evolution. Depending on the option this can be improved. 


## Algorithm 2.2
To find the price of a path-independent option

2. Compute the payouts at maturity (the leaves of the lattice). $V_T^j$
```{r, tidy=TRUE, eval=FALSE, highlight=FALSE }
for (j in 1:T) {
  V[T][j] = max((S[T][j]), 0)## for European call for instance
}
```

3. Compute $V[t][j]$ backwards in time until you get to $V[0][0]$:
```{r, tidy=TRUE, eval=FALSE, highlight=FALSE }
# d, u are down and up steps
algo21 <- function(u,d,r, V, T){
  q_u = (e^r-d)/(u-d) # risk-free probability of up
  q_d = 1-q_u
  for (t in T-1:0){
    for (j in 0:t){
      V[t][j] = (e^(-r)*(q_u* V[t+1][j+1] + q_d*V[t+1][j])
    }
  }
  return V[0][0]
}
```


## American Option pricing
In american options we can exercise at any time up until maturity. 
We must keep track of two values at time $t$: 

1. $V_t^{ex}$: The current exercise value, payout at time $t$. 
2. $V_t^{cont}$: The continuation value, the value of the option if not exercised at time t (present value of what it is expected to give us). 

Assuming rational agents, at any time $t$ in an American option, we exercise if $V_t^{ex} > V_t^{cont}$ and hold otherwise. 

We can adapt **Algorithm 2.1** to price this:

### Algorithm 2.3:
1. Compute all possible payouts at maturity $V_T(\vec w)$ corresponsing to all possible states $\vec w = (\omega_0, \dots, \omega_T) \in \Omega=\{up, down\}$ 
2. Go backwards in time. For $n= N-1, \dots, 0$ do: 
  * for all states $\vec w$ do:
      a. Compute the continuation value $V_n^{cont} (\vec w)= e^{-r \Delta t} (q^u V_{n+1}(\vec w, up) + q^dV_{n+1}(\vec w, down))$
      b. Compute the execution value $V_n^{ex} (\vec w) = max\{S_n(\vec w)-K, 0 \}$
      c. Compute the rational value $V_n(\vec w) = max\{ V_n^{ex} (\vec w), V_n^{cont} (\vec w)\}$
      
3. Return $V_0$

Just like with Algorithm 2.2, if we have a path-independent option we can simplify this algorithm by expressing values in terms of number of up states $V_t^j$ as opposed to expressing them in terms of entire state evolutions $V_t(\vec \omega)$

\newpage

# Black-Scholes Model
We assume the stock price $S_t$ is an Ito-process. Ito-processes satisfy Geometric Brownian Motion dynamics. 

\[ dS_t = \mu S_t dt+ \sigma S_t dB_t\]

This line encapsulates alll our asusmptions. 

Also we can show using Ito's lemma that the natural logarithm of an ito process satisfies Arithmetic Brownian motion dynamics: 

\[ dln(S_t) = (\mu -\frac{\sigma^2}{2})dt+ \sigma dB_t \] 

And we can solve this without Ito Integrals because the coefficients of the infinitesimal changes ($dt, dB_t$) are constant. 

The solution is $S_t = S_0 \exp\{(\mu -\frac{\sigma^2}{2})t+ \sigma B_t\}$

For Monte Carlo Algorithms, we will use the aforementioned GBM solution as well as the fact that: 
\[V_0 = e^{rT} \mathbb{E}^{\mathbb{Q}}(\text{payout}((S_t)_{t\in [0,T]})) \]

Also under $\mathbb{Q}$, $\mu = r$ in our GBM solution (**Not sure how to prove this**), so we can say: 
\[ S_t = S_0 \exp\{(r -\frac{\sigma^2}{2})t+ \sigma B_t\} \]



### Monte Carlo Algorithms
Simulation algorithms in which we sample from a distribution. 
We will sample $n$ different paths and then compute $n-$different payouts based on the realizations. We will then return the average. 

## Algorithm 3.1
Given $n$: number of simulation runs, $N:$ number of time discretizations suth that for a maturity time horizon $T$, $T/N = \Delta t$. 

1. Discretize time s.t. we only simulate at $0 < t_1 < t_2 < \dots < t_N$
2. For $i = 1, \dots n$ (number of simulation runs):
    * Sample a discretized path $S_1, \dots S_N$ corresponding to $t_1, \dots t_N$.
    * Compute the payout in simulation run $i$, say $V_N^j$. Unlike with binomial models we don't need to know $V_t$ for $t<N$. 
3. Return the estimate $\hat V_0 = e^{-rT}\frac{1}{n}V^i_N$ (the discounted expected value of the average payout at maturity.)

This can be used for path-dependent (Asian) or independent options (European, rainbow?) but, as the algorithm is, the options should be exercised at maturity i.e. no american-type options (you probably just modify it a bit like with algo 2.3)

### Sources of error for Monte Carlo: 
* Discretization error as we sample only at $t_1 < t_2 < \dots < t_N$ and not continuosuly (impossible on a finite resource computer)
* Sampling error or MC error, due to only sampling $n$ paths. We can't ever sample all uncountably many paths. 



