---
output: pdf_document
---

# Q1) Monte Carlo Algorithm for Rainbow Option 

## a) 
For $d-$correlated assets the prices at maturity are $(S^{(j)}_T)_{i=1, \dots, d}$.
Under $\mathbb{Q}$, $T=1$, and the Black Scholes model, the solution to the price at maturity for all $j=1,2,\dots,d$ is: 

\[S_T^{(j)} = S_0^{(j)} \exp(r-(\sigma^{(j)})^2/2 + \sigma^{(j)} W_T^{(j)}\]

Where for each stock $j =1,2,\dots, d$:

  * $S_0^{(j)}$ is its initial price of stock
  * $\sigma^{(j)}$ is its standard deviation
  * $W_T^{(j)}$ is is the brownian motion at maturity such that for another stock $i$ we have correlated brownian motion processes $Cor(W_t^{(j)}, W_t^{(i)})=\rho$. 
  
Define the correlation matrix.
\[
\Sigma = 
\begin{bmatrix} 
    1 & \rho  & \dots & \rho \\
    \rho & 1  & \dots & \rho \\
    \vdots & \vdots &  \ddots &\vdots \\
    \rho  &  \rho & \dots &1
\end{bmatrix}
    \in\mathbb{R}^{d\times d}
\]

We want $X =  (W_T^{(1)}, \dots , W_T^{(d)}) \sim N_d(0, \Sigma)$, so let $Z^j \sim N(0,1)$ for $j=1,\dots d$ such that for a matrix $A \in \mathbb{R}^{d\times d}$ satisfying $\Sigma = AA^T$ we have $X = AZ$ (equal in distribution). Let $A$ be the **Cholesky factor** of $\Sigma$.

*****

**Algorithm**

1. Given $S_0^{(j)}, \sigma^{(j)}$ for all assets $j=1,\dots, d$
2. Initialize $V_T :: \text{float}[d] \gets 0 \in \mathbb{R}^d$
3. Compute $A \gets \text{cholesky}(\Sigma)$
4. for $i = 1, \dots, n$ do:
    * Sample $Z_1, \dots, Z_d \sim^{iid} N(0,1)$ and let $Z = (Z_1, \dots, Z_d)$
    * Let $X = AZ \in \mathbb{R}^d$ (matrix multiplication)
    * Let $S_{MAX} = -1$
    * for $j=1, \dots , d$:
      - Let $S_T^{(j)} = S_0^{(j)} exp\{r-(\sigma^{(j)})^2/2 + \sigma^{(j)} X[j] \}$
      - If $(S_T^{(j)} > S_{MAX})$ : $S_{MAX} = S_T^{(j)}$ 
    * Set $V_T[i] = \exp(-r) \times \max(S_{MAX}-K, 0)$
5. return $\text{mean}(V_T) \pm \frac{1.96\times \text{sd}(V_T)}{\sqrt{n}}$



\newpage 

## b) 

```{r}
getAmatrix <-function(rho, d){
  ## Construct correlation matrix Sigma for d-RV's with pairwise correlation rho. 
  Sigma <- matrix(rep(rho, d^2), nrow = d, ncol = d) # populate with rho
  diag(Sigma) <- 1 # Set diagonal equal to 1
  ## Get matrix A such that AA^T = Sigma
  t(chol(Sigma))
}

partA <- function(d, S0s, sigmas, rho, K, r, n) {
  A <- getAmatrix(rho, d)
  vals <- rep(0, n) # values at maturity for each simulation
  for (i in 1:n){
    Z <- rnorm(d)
    X <- A%*%Z
    Smax <- -1
    for (j in 1:d){
      STj <- S0s[j] * exp((r-sigmas[j]^2/2) + sigmas[j]*X[j])
      if (STj > Smax) {
        Smax = STj
      }
    }
    vals[i] <- max(exp(-r)*(Smax-K), 0)
  }
  MCest <- mean(vals)
  MCsd <- sd(vals)
  alpha <- qnorm(.975)
  lo <- MCest-alpha*MCsd/sqrt(n)
  hi <- MCest+alpha*MCsd/sqrt(n)
  ans <- c(MCest, lo, hi)
  ## names(ans) <- c("estimate", "low", "high")
  return(ans)
}

partB <- function(rho) {
  set.seed(123)
  ans<-partA(d= 2, S0s = c(100, 100), sigmas = c(.2, .3), rho=rho, K=110, r=0.05, n=100000)
  return(ans[1])
}

```
\newpage 

```{r}
rhos <- seq(-0.9, 0.9, 0.1)
estimates <- sapply(rhos, partB)
df <- data.frame(rhos, estimates)
print(df)
plot(x = rhos, y = estimates, main = "Correlation vs MC price estimates", 
     ylab = "estimate", xlab = "pairwise correlation")

```












